---
title: "NBA Analysis"
author: "Marco Lagos"
date: "2023-04-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(stringr)
```

Read in the data sets (advanced_stats.csv, nba_salaries.csv, nba_team_wins.csv, 
shooting_opponent.csv, shooting_team.csv, total_opponent.csv, total_team.csv)
```{r}
nba_salaries <- read.csv("./data/nba_salaries.csv")
nba_team_wins <- read.csv("./data/nba_team_wins.csv")

advanced_stats <- read.csv("./data/advanced_stats.csv")

shooting_stats <- read.csv("./data/shooting_team.csv")
opponent_shooting_stats <- read.csv("./data/shooting_opponent.csv")

total_stats <- read.csv("./data/total_team.csv")
opponent_total_stats <- read.csv("./data/total_opponent.csv")

gdp_deflator <- read.csv("./data/gdp_deflator_by_year.csv")

# View(nba_salaries)
# View(nba_team_wins)
# View(advanced_stats)
# View(shooting_stats)
# View(opponent_shooting_stats)
# View(total_stats)
# View(opponent_total_stats)
# View(gdp_deflator)
```

Linear regression plot of team wins on total team salary. There needs to be 
inflation adjustment to the previous years so that salary is consistent. Then,
we will plot the the team wins to the total adjusted team salary by year.

We can use different index to make our USD adjustments over the course of the 20
years. Let us just choose our base year to be 2020.
1. Consumer Price Index (CPI)
	- Base Year CPI * (Current Year value / Current Year CPI)
	- This will give the adjusted amount of USD for that current year according
	  to the CPI index
2. Gross Domestic Product Deflator (GDP Deflator)
	- Base Year GDP * (Current Year value / Current Year GDP)
	- This will give the adjusted amount of USD for that current year according
	  to the CPI index
Why did we choose 2020 as the base year? Our intentions are to adjust the USD as
close as we can for the current year, so that viewers can better understand our 
analysis.

```{r}
adjust_salary_gdp_deflator <- function(base_year, nba_salaries, gdp_deflator) {
  base_year_gdp_value <- filter(gdp_deflator, year == base_year)$value

  adjusted_salary <- c()
  for (i in 1:nrow(nba_salaries)) {
  	current_year <- nba_salaries[i, 6]
	current_year_gdp_value <- filter(gdp_deflator, year == current_year)$value
	factor <- base_year_gdp_value / current_year_gdp_value
	adjusted_salary[i] <- nba_salaries[i, 5] * factor
  }
  nba_salaries <- cbind(nba_salaries, adjusted_salary)
  
  return(nba_salaries)
}

base_year <- 2020
adjusted_nba_salaries <- adjust_salary_gdp_deflator(base_year, nba_salaries, gdp_deflator)

total_salary <- adjusted_nba_salaries %>% group_by(year, team) %>% summarize(total_salary = sum(adjusted_salary))

# colnames(total_salary) is "year", "team", "total_salary"

create_total_salary_to_wins <- function(total_salary, wins) {
	wins_long <- nba_team_wins %>% pivot_longer(cols = names(nba_team_wins)[-1], names_to="team", values_to = "wins") %>% mutate(team = str_replace_all(team, "\\.", " "))
	
	total_salary$team <- str_to_lower(str_trim(total_salary$team))
	wins_long$team <- str_to_lower(str_trim(wins_long$team))
	
	merged_data <- left_join(total_salary, wins_long, by = c("year", "team"))
	names(merged_data) <- c("year", "team", "total_salary", "wins")
	return(merged_data)
}
total_salary_to_wins <- create_total_salary_to_wins(total_salary, nba_team_wins)
# colnames(total_salary_to_wins) is "year", "team", "total_salary", and "wins"
total_salary_to_wins <- na.omit(total_salary_to_wins)

# total_salary column for 2000, Atlanta Hawks was 23346098.
# Total wins for 2000, Atlanta Hawks was 25.

```
```{r}

# Get rid of the asterisks in every Team Name in advanced_stats.csv
advanced_stats <- read.csv("./data/advanced_stats.csv")
advanced_stats
advanced_stats$team <- stringr::str_replace(advanced_stats$team, '\\*', '')
advanced_stats

# names(advanced_stats)[] = "new_column_name"

# append = FALSE will destroy the existing file and then replace it with the variable
# advanced_stats which we are writing to that file.
# quote = FALSE will make sure that no elements are surrounded by quotation marks
# when we write these elements to the file.
# write.csv(advanced_stats, "./data/advanced_stats_without_asterisk.csv", append = FALSE, quote = FALSE)

# NOTE: The "Unnamed: 17", "Unnamed: 22", and "Unnamed: 27" columns of advanced_stats are
# not useful. They contain
# empty values or no values (idk). Just do not use these three columns.


create_wins_long <- function(wins) {
  # NOTE THAT THIS FUNCTION USES nba_team_wins, I THINK IT SHOULD TECHNICALLY USE "wins"
  # which is more specific to what was actually passed into this function create_wins_long.
	wins_long <- nba_team_wins %>% pivot_longer(cols = names(nba_team_wins)[-1], names_to="team", values_to = "wins") %>% mutate(team = str_replace_all(team, "\\.", " "))
	
	wins_long$team <- str_to_lower(str_trim(wins_long$team))
	
	return(wins_long)
}

# nba_team_wins
wins_long_res <- create_wins_long(nba_team_wins)
wins_long_res

# Assume that every string in the "team" column of wins_long_inside
# has already been trimmed and converted to lower case, BEFORE you pass it
# into this function.
create_advanced_stats_to_wins <- function(advanced_stats_inside, wins_long_inside) {
	
  # All the team names in wins_long$team is already trimmed and converted to lower case.
  # Therefore we should trim and convert to lower case all the team names inside
  # advanced_stats_inside.
	advanced_stats_inside$team <- str_to_lower(str_trim(advanced_stats_inside$team))
	
	# advanced_stats_bool_of_team_vec <- ifelse(wins_long_inside.contains(dvanced_stats_inside$team))
	# advanced_stats_bool_of_year_vec <- ifelse(wins_long_inside.contains(dvanced_stats_inside$team))
	
	
	
	# MAYBE USE THIS FOR LATER.
	# advanced_stats_bool_of_team_vec <- ifelse(wins_long_inside.contains(advanced_stats_inside$team)
	#                                          AND wins_long_inside.contains(advanced_stats_inside$year))
	
	# advanced_stats_inside <- advanced_stats_inside[advanced_stats_bool_of_team_vec]
	
	
	#	wins_long_bool_of_team_vec <- ifelse(advanced_stats_inside.contains(wins_long_inside$team)
	#                                          AND advanced_stats_inside.contains(wins_long_inside$year))
	
	# wins_long_inside <- wins_long_inside[wins_long_bool_of_team_vec]
	
	
	
	# We could do a simple ifelse or a filter that removes all the rows in advanced_stats_inside
	# and removed all the rows in wins_long_inside that correspond to a year-team combination
	# that does NOT exist in both advanced_stats_inside and wins_long_inside.
	
	# left_join has a argument called "fill = list(wins = 0)"
	merged_data <- left_join(advanced_stats_inside, wins_long_inside, by = c("year", "team"))
	# names(merged_data) <- c("year", "team", "total_salary", "wins")
	return(merged_data)
}

advanced_stats_to_wins <- create_advanced_stats_to_wins(advanced_stats, wins_long_res)
advanced_stats_to_wins
```
Making the plot of team wins over total salary paid to players for that year. The
salaries have been adjusted and any NULL wins have been removed.
```{r}
x <- total_salary_to_wins$total_salary
y <- total_salary_to_wins$wins

model <- lm(y ~ x)
summary(model)
plot(x, y)




```

