---
title: "NBA Analysis"
author: "Marco Lagos"
date: "2023-04-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in the data sets (advanced_stats.csv, nba_salaries.csv, nba_team_wins.csv, 
shooting_opponent.csv, shooting_team.csv, total_opponent.csv, total_team.csv)
```{r}
nba_salaries <- read.csv("./data/nba_salaries.csv")
nba_team_wins <- read.csv("./data/nba_team_wins.csv")

advanced_stats <- read.csv("./data/advanced_stats.csv")

shooting_stats <- read.csv("./data/shooting_team.csv")
opponent_shooting_stats <- read.csv("./data/shooting_opponent.csv")

total_stats <- read.csv("./data/total_team.csv")
opponent_total_stats <- read.csv("./data/total_opponent.csv")

View(nba_salaries)
View(nba_team_wins)
View(advanced_stats)
View(shooting_stats)
View(opponent_shooting_stats)
View(total_stats)
View(opponent_total_stats)
```

Linear regression plot of team wins on total team salary. There needs to be 
inflation adjustment to the previous years so that salary is consistent. Then,
we will plot the the team wins to the total adjusted team salary by year.
```{r}
```


