---
title: "Preprocessing Data"
author: "Marco Lagos"
date: "2023-04-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(stringr)
```

Read in the data sets (advanced_stats.csv, nba_salaries.csv, nba_team_wins.csv, 
shooting_opponent.csv, shooting_team.csv, total_opponent.csv, total_team.csv)
```{r}
nba_salaries <- read.csv("./data/nba_salaries.csv")
nba_team_wins <- read.csv("./data/nba_team_wins.csv")

advanced_stats <- read.csv("./data/advanced_stats.csv")

shooting_stats <- read.csv("./data/shooting_stats.csv")
opponent_shooting_stats <- read.csv("./data/opponent_shooting_stats.csv")

total_stats <- read.csv("./data/total_stats.csv")
opponent_total_stats <- read.csv("./data/opponent_total_stats.csv")

gdp_by_year <- read.csv("./data/gdp_by_year.csv")
cpi_by_year <- read.csv("./data/cpi_by_year.csv")

# View(nba_salaries)
# View(nba_team_wins)
# View(advanced_stats)
# View(shooting_stats)
# View(opponent_shooting_stats)
# View(total_stats)
# View(opponent_total_stats)
# View(gdp_deflator)

clean_data <- function(data, path) {
	# remove NA columns
	cols_to_remove <- names(data)[grepl("^unnamed", names(data))]
	data <- data[, !names(data) %in% cols_to_remove]
	
	# trim white space and make lowercase all team names
	if ("team" %in% names(data)) {
		data$team <- tolower(trimws(data$team))
		data$team <- gsub("\\*$", "", data$team)
		data %>% mutate(team = str_replace_all(team, "\\.", " "))
	}
	write.csv(data, path, row.names = FALSE, quote = FALSE)
	return(data)
}

nba_salaries <- clean_data(nba_salaries, "./data/cnba_salaries.csv")

nba_team_wins <- 
	nba_team_wins %>% 
	pivot_longer(cols = names(nba_team_wins)[-1], names_to="team", values_to = "wins") %>%
	mutate(team = str_replace_all(team, "\\.", " "))
nba_team_wins <- clean_data(nba_team_wins, "./data/nba_team_wins.csv")

advanced_stats <- clean_data(advanced_stats, "./data/advanced_stats.csv")
shooting_stats <- clean_data(shooting_stats, "./data/shooting_stats.csv")
opponent_shooting_stats <- clean_data(opponent_shooting_stats, "./data/opponent_shooting_stats.csv")
total_stats <- clean_data(total_stats, "./data/total_stats.csv")
opponent_total_stats <- clean_data(opponent_total_stats, "./data/opponent_total_stats.csv")
gdp_by_year <- clean_data(gdp_by_year, "./data/gdp_by_year.csv")
cpi_by_year <- clean_data(cpi_by_year, "./data/cpi_by_year.csv")
```


```{r}
total_salary <- 
	adjusted_nba_salaries %>% 
	group_by(year, team) %>% 
	summarize(total_salary = sum(adjusted_salary))
write.csv(total_salary, "./data/total_salary_by_team.csv", row.names = FALSE, quote = FALSE)

create_stat_to_wins <- function(stat) {
	merged_data <- left_join(stat, nba_team_wins, by = c("year", "team"))
	return(merged_data)
}
```
Linear regression plot of team wins on total team salary. There needs to be 
inflation adjustment to the previous years so that salary is consistent. Then,
we will plot the the team wins to the total adjusted team salary by year.

We can use different index to make our USD adjustments over the course of the 20
years. Let us just choose our base year to be 2020.
1. Consumer Price Index (CPI)
	- Base Year CPI * (Current Year value / Current Year CPI)
	- This will give the adjusted amount of USD for that current year according
	  to the CPI index
2. Gross Domestic Product Deflator (GDP Deflator)
	- Base Year GDP * (Current Year value / Current Year GDP)
	- This will give the adjusted amount of USD for that current year according
	  to the CPI index
Why did we choose 2020 as the base year? Our intentions are to adjust the USD as
close as we can for the current year, so that viewers can better understand our 
analysis.

Why did we choose CPI? The GDP deflator is a measure of changes in the domestic 
price of production over time and includes government purchases and business 
investment as it comes from total GDP. The CPI is just consumption and is more 
geared towards household purchases of goods and services. That is more relevant 
for a person’s salary.

```{r}
adjust_salary_cpi <- function(base_year, salaries, adjuster) {
  base_year_value <- filter(adjuster, year == base_year)$value

  adjusted_salary <- c()
  for (i in 1:nrow(salaries)) {
  	current_year <- salaries[i, 6]
	current_year_value <- filter(cpi, year == current_year)$value
	factor <- base_year_value / current_year_value
	adjusted_salary[i] <- salaries[i, 5] * factor
  }
  salaries <- cbind(salaries, adjusted_salary)
  
  return(salaries)
}

base_year <- 2020
adjusted_nba_salaries <- adjust_salary_cpi(base_year, nba_salaries, gdp_deflator)
write.csv(adjusted_nba_salaries, "./data/adjusted_nba_salaries.csv", row.names = FALSE, quote = FALSE)
```
